---
title: "Homework 4"
author: "Austin Dollar"
date: "9/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

The written out answer to question 1 is included in the scanned pdf attached in the homework submission


## Question 2

Generate data from the gamma distribution Γ(shape = 3, rate = 6), and then use optim to estimate shape and rate via the log-likelihood function for the gamma
distribution. Note that you can use the log-likelihood function you found as part of
problem 1.

```{r Q2}

ll.gamma <- function(theta, x){
  
  a <- theta[1]
  B <- theta[2]
  n <- length(x)
  
    (n*log(B))-(n*log(gamma(a)))+(a-1)*sum(log(x)) - sum(B*x)
    
    
} 

data <- rgamma(1e5, shape = 3, rate = 6)


optim(par=c(7,6), fn=ll.gamma, method="L-BFGS-B", lower=c(1e-6,1e-6),control=list(fnscale=-1), 
      x = data)

```

As you can see above, optim returns the expected value for the rate, as par = 6 as well as the shape, whereas par = 3. 


