---
title: "Assignment 1"
author: "Austin Dollar"
date: "9/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rlist)
```

## Question 1


```{r Q1}

#func definition

sum_mults <- function(nums, n)
{
  f <- 0
  
  for(x in 1:n-1)
  {
    for(i in nums)
    {

      if(x %% i == 0)
      {
        f <- f+x
        break
      }
       
    }
  }
  return(f)
  
}

#test function
Q1_1<- sum_mults(c(3,5), 30)
Q1_2<- sum_mults(c(2,8), 40)
Q1_3<- sum_mults(c(4,6), 50)


```

Test of sum_mults(c(3,5), 30) = `r Q1_1`

Test of sum_mults(c(2,8), 40) = `r Q1_2`

Test of sum_mults(c(4,6), 50) = `r Q1_3`



## Question 2


```{r Q2}

collatz_len <- function(n)
{
  t <-0
  
  while(n > 1)
  {
    if (n%%2 == 0)
    {
      
      n = n/2
      t = t+1
      
    }else
    {
      n = 3*n+1
      t = t+1
    }
    
  }
  
  t = t+1
  
  return(t)
  
}


collat_1 <- collatz_len(17)
collat_2 <- collatz_len(15)
collat_3 <- collatz_len(21)


```

Test of collatz_len(17) = `r collat_1`

Test of collatz_len(15) = `r collat_2`

Test of collatz_len(21) = `r collat_3`


## Question 3


```{r Q3}

reverse<- function(v)
{
  return(rev(v))
}

reversal_1 <- reverse(c(1,2,3))
reversal_2 <- reverse(c(2,2,3))
reversal_3 <- reverse(c(4,5,6))

```

Test of reverse(c(1,2,3)) = `r reversal_1`

Test of reverse(c(2,2,3)) = `r reversal_2`

Test of reverse(c(4,5,6)) = `r reversal_3`


## Question 4


```{r Q4}

drop <- function(v, n)
{
  c <- 0
  
  v1 <- character()
  
  for(i in v)
  {
    c <- c + 1
    
    if(c == n)
    {
      v1 <- c(v1, v[i])
      
      c = 0
    }
  }
  
  v <- setdiff(v, v1)
  return(v)
}


dropper_1 <- drop(c(1,2,3,4,5), 2)
dropper_2 <- drop(c(1,2,3,4,5), 4)
dropper_3 <- drop(c(1,2,3,4,5), 3)

```

Test of drop(c(1,2,3,4,5), 2) = `r dropper_1`

Test of drop(c(1,2,3,4,5), 4) = `r dropper_2`

Test of drop(c(1,2,3,4,5), 3) = `r dropper_3`


## Question 5


```{r Q5}

intersect_3 <- function(v, w, x)
{
  v1 <- intersect(v,w)
  v1 <- intersect(v1, x)
  
  return(v1)
  
}

inter <- intersect_3(c(1,2,3,1), c(1,1,3,2), c(3,1,9,1))
inter_1 <- intersect_3(c(1,2,3), c(3,4,5), c(1,2,3,4))
inter_2 <- intersect_3(c(4,5,6,7), c(3,4,5), c(5,7,8))

```

Test of intersect_3(c(1,2,3,1), c(1,1,3,2), c(3,1,9,1)) = `r inter`

Test of intersect_3(c(1,2,3), c(3,4,5), c(1,2,3,4)) = `r inter_1`

Test of intersect_3(c(4,5,6,7), c(3,4,5), c(5,7,8)) = `r inter_2`


## Question 6


```{r Q6}

filter_vec <- function(v,p)
{
  v1 <- integer()
  
  for (i in v)
  {
    if(p(v[i]) == TRUE)
    {
      v1 <- c(v1, v[i])
    }
  }
  
  return(v1)
}

p <- function(x){return(x>3)}
l <- 1:6
m <- filter_vec(l,p)



```

Test of filter_vec(l,p) = `r m`

## Question 7


```{r Q7}

n_fibs <- function(n)
{
  v <- integer()
  
  for(i in 1:n)
  {
    if(i == 1 | i == 2)
    {
      
      v <- c(v, 1)
      
    }else
    {
      cur <- v[i-1] + v[i-2]
      v <- c(v, cur)
      
    }
    
  }
  
  return(v)
}

fib <- n_fibs(2)
fib_1 <- n_fibs(3)
fib_2 <- n_fibs(8)

```

Test of n_fibs(2) = `r fib`

Test of n_fibs(3) = `r fib_1`

Test of n_fibs(8) = `r fib_2`


## Question 8


```{r Q8}

shift <- function(v,n)
{

  if (n == 0)
  {
    
    return(v)
    
  }else if(n > 0)
    {
      return(c(tail(v, -n), head(v, n)))
  }
  else
  {
    return(c(tail(v, n), head(v, -n)))
  }
  
  
}

shifter1 <- shift(c(1,2,3,4), 2)
shifter2 <- shift(c(1,2,3,4), -3)
shifter3 <- shift(c(1,2,3,4), -1)

```

Test of shift(c(1,2,3,4), 2) = `r shifter1`

Test of shift(c(1,2,3,4), -3) = `r shifter2`

Test of shift(c(1,2,3,4), -1) = `r shifter3`


## Question 9


```{r Q9}

rem_consec_dups <- function(v)
{
  v1 = integer()
  k <- 99999
  
  for(i in 1:length(v))
  {
    if(v[i] != k)
    {
      v1 <- c(v1, v[i])
    }
    
    k <- v[i] 
  }
  return(v1)
}

remmer <- rem_consec_dups(c(1,1,1,2,3,3,1,2,2))
remmer_1 <- rem_consec_dups(c(1,1,1,2,3,3,1,2,2))
remmer_2 <- rem_consec_dups(c(1,1,1,2,3,3,1,2,2))

```

Test of rem_consec_dups(c(1,1,1,2,3,3,1,2,2)) = `r remmer`

Test of rem_consec_dups(c(11,2,2,3,3,1,2)) = `r remmer_1`

Test of rem_consec_dups(c(1,1,1,2,3,3,1,2,24,5,5,6,6,6,7)) = `r remmer_2`


## Question 10


```{r Q10}

n_even_fibs <- function(n)
{
  v <- integer()
  
  for(i in 1:30)
  {
    if(i == 1 | i == 2)
    {
      
      v <- c(v, 1)
      
    }else
    {
      cur <- v[i-1] + v[i-2]
      v <- c(v, cur)
      
    }
    
  }
  
  fiblist <- list(name = 1)
  
for(x in 1:length(v))
{
  if(v[x] %% 2 == 0)
  {
    zz = as.character(x)
    list.append(fiblist, zz = v[x])
  }
}
  print(fiblist)
}

efibs <- n_even_fibs(5)

```

