---
title: "Assignment 2"
author: "Austin Dollar"
date: "9/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Q1 

Suppose you have a vector of indices of the months from 2020 to
2021 (let’s call it time), and that you have two other vectors of
indices summer2020 and summer2021, which store the indices of
the summer months from their respective years. Use set operations
to find the indices of the non-summer months.


```{r Q1}

time <- 1:24
summer2020 <- c(7, 8, 9)
summer2021 <- c(19, 20, 21) 
summer <- union(summer2020, summer2021)
nonsummer <- setdiff(time, summer)

```

The indices of non summer months are `r nonsummer`

## Q2

At a county fair carnival game, there are 25 balloons on a board, 8
of which are yellow, 10 red, and 7 green. Given that the first
balloon hit is red, what is the probability that the next balloon hit
will also be red? Hence, find P(2red |1red ). Additionally, what is
the combined probability of both hit balloons being red?


```{r Q2}

# P(first_red) = 10/25

# P(first_red | second_red) = 9/24
#Probability that second balloon is red

# P(first_red & second-red) = (10*9)/(25*24) = 0.15
# Probability that they are both red


```


## Q3

Roll a six-sided die once. Let E1 = {5, 6} and E2 = {6}. Are E1
and E2 independent?


```{r, Q3}

# E1 = {5,6}   
# E2 = {6}   

p.E2 = 1/6
p.E1 = 2/6

p.E1andE2 = 1/6
p.E1andE2 ==  p.E1*p.E2

```

E1 and E2 are not independent because, as shown above, the product of the probabilities is not the same as their intersection.


## Q4 

Roll a six-sided die once. Let E1 = {2, 3} and E2 = {2, 4, 6}. Are
E1 and E2 independent?


```{r, Q4}

#E1 = {2,3}
#E2 = {2,4,6}

pp.E1 = 2/6
pp.E2 = 3/6

pp.E1andE2 = 1/6
pp.E1andE2 == pp.E1*pp.E2

```

E1 and E2 are independent because, as shown above, the product of the probabilities is the same as their intersection.


## Q5 

The three components of the parallel system, D, E, and F, will
function with probabilities pD = 0.92, pE = 0.88, and pF = 0.85,
independently of each other. What is the probability the parallel
system will function?


```{r, Q5}

pD <- 0.92
pE <- 0.88
pF <- 0.85

pDf <- 1 - pD
pEf <- 1 - pE
pFf <- 1 - pF

pFail <- pDf*pEf*pFf
pFunction = 1 - pFail

```

The probability the system will function is `r pFunction`

## Q6

Consider the data set arules::Groceries. Also import the CSV
version from the listed URL (after removing the space).


```{r, 6Question}

suppressMessages(library(arules))

data(Groceries)

arGroceries <- apriori(Groceries,
          parameter=list(supp=0.001, conf=0.8))

inspect(head(arGroceries, n=4))

# Import from CSV version from URL
url <- "https://raw.githubusercontent.com/roualdes/data/master/groceries.csv"

groceries <- read.csv(url)

```


Calculate the support, confidence, and lift for row 3 and row 4
using the rules developed in the Lecture 02 notes. Use mean(), &,
and | to do so, and use complete sentences to interpret the
numbers you get. Check your work with arGroceries.


```{r, Q6}

p.A <- mean(groceries$whole.milk)  #same for both rows

# row 3 
p.B <- mean(groceries$yogurt & groceries$cereals)

row3.supp = mean(groceries$whole.milk & groceries$yogurt & groceries$cereals)

row3.conf = row3.supp/p.B

row3.lift = row3.conf/p.A


#row4
p.B <- mean(groceries$butter & groceries$jam)

row4.supp = mean(groceries$whole.milk & groceries$butter & groceries$jam)

row4.conf = row4.supp/p.B

row4.lift = row4.conf/p.A

```

The Support value that was found for row 3 is `r row3.supp``, and matches with the sample code. 

The calculated confidence value for row 3 is `r row3.conf`` and matches our sample value.

lift was calculated for row 3 at `r row3.lift`` and matches what the answer key says. 


The Support value that was found for row 4 is `r row4.supp``, and matches with the sample code. 

The calculated confidence value for row 4 is `r row4.conf`` and matches our sample value.

lift was calculated for row 4 at `r row4.lift`` and matches what the answer key says. 























